---
title: "COVID-19-Austria: Evidence of spread slowing down"
author: "Thomas.Strehl@at.ibm.com"
date: "Mai 03, 2020"
geometry: "left=1.5cm,right=1cm,top=1.5cm,bottom=1.5cm"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    fig_width: 7
    fig_height: 4
    fig_caption: false
    number_sections: true     
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, echo=FALSE}
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
```

\newpage

# Status 

## Prepare data
```{r}
csvFile <- "/home/at062084/DataEngineering/COVID-19/COVID-19-Austria/bmsgpk/data/COVID-19-austria.states.curated.status.csv"
dv <- read.csv(csvFile, stringsAsFactors=FALSE) %>% 
  dplyr::mutate(Stamp=as.POSIXct(Stamp), Date=date(Stamp)) 

df <- dv %>% 
  dplyr::arrange(Stamp) %>%
  dplyr::group_by(Date,Status) %>%
  dplyr::filter(row_number()==n()) %>%
  dplyr::ungroup() %>%
  dplyr::select(Date,Status,Count=AT) %>%
  tidyr::spread(key=Status, val=Count) %>%
  dplyr::mutate(newConfirmed=Confirmed-lag(Confirmed),newDeaths=Deaths-lag(Deaths)) %>%
  dplyr::filter(!is.na(newConfirmed), newConfirmed>0) %>%
  dplyr::select(1,8,2,9,3:7)
tail(df,12)

# Vienna:
dw <- dv %>% 
  dplyr::arrange(Stamp) %>%
  dplyr::group_by(Date,Status) %>%
  dplyr::filter(row_number()==n()) %>%
  dplyr::ungroup() %>%
  dplyr::select(Date,Status,Count=W) %>%
  tidyr::spread(key=Status, val=Count) %>%
  dplyr::mutate(newConfirmed=Confirmed-lag(Confirmed),newDeaths=Deaths-lag(Deaths)) %>%
  dplyr::filter(!is.na(newConfirmed), newConfirmed>0) %>%
  dplyr::select(1,8,2,9,3:7)
tail(dw,12)
```


## Status ~ Stamp

### Status ~ Stamp (all)

```{r}
ggplot(data=df %>% tidyr::gather(key=Status,val=Count, 2:8), aes(x=Date, y=Count, colour=Status, shape=Status)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous(trans = 'log10') +
  ggtitle("x=Date, y=log10(Count)")
```

### Status ~ Stamp (selected)

```{r}
ggplot(data=df %>% tidyr::gather(key=Status,val=Count, 2:8) %>% dplyr::filter(Status!="Tested", Status !="Confirmed", Status!="Recovered"), 
       aes(x=Date, y=Count, colour=Status, shape=Status)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous() +
  ggtitle("x=Date, y=Count")
```

### Confirmed ~ Stamp
```{r}
ggplot(data=df, aes(x=Date, y=Confirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.33, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous(breaks=0:13*1000) +
  ggtitle("x=Date, y=Confirmed")
```

## Tested
### Confirmed/Tested ~ Tested

```{r}
ggplot(data=df, aes(x=Tested, y=Confirmed/Tested)) + 
  geom_point() + geom_line(linetype=3) + 
  scale_x_continuous(breaks=seq(0,300000,by=20000)) +
#  scale_y_continuous(trans = 'log10') +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  ggtitle("x=Tested, y=Confirmed/Tested")
```

## newConfirmed
### newConfirmed ~ Stamp
```{r}
ggplot(data=df, aes(x=Date, y=newConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.25, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  ggtitle("x=Date, y=newConfirmed")
```

### newConfirmed ~ Confirmed
```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  scale_x_continuous(breaks=0:100*1000) +
  stat_smooth(span=0.33, se=FALSE, size=.33) +
  ggtitle("x=Confirmed, y=newConfirmed")
```

### newConfirmed ~ log10(Confirmed)
```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed)) + 
  scale_x_continuous(trans = 'log10', breaks=c(10,20,50,100,200,500,1000,2000,5000,10000)) +
#  scale_y_continuous(trans = 'log10') +
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.33, se=FALSE, size=.5) +
  ggtitle("x=log10(Confirmed), y=newConfirmed")
```



## newConfirmed/Confirmed
### newConfirmed/Confirmed ~ Date
```{r}
ggplot(data=df, aes(x=Date, y=newConfirmed/Confirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  ggtitle("x=Date, y=newConfirmed/Confirmed")
```

### newConfirmed/Confirmed ~ Confirmed
```{r}
ggplot(data=df, aes(x=(Confirmed), y=newConfirmed/Confirmed)) + 
  geom_point() + geom_line(linetype=3) +
   stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_continuous(breaks=0:100*1000) +
  ggtitle("x=Confirmed, y=newConfirmed/Confirmed")
```

### newConfirmed/Confirmed ~ log10(Confirmed)
```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed/Confirmed)) + 
  scale_x_continuous(trans = 'log10') +
  #scale_y_continuous(trans = 'log10') +
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  ggtitle("x=log10(Confirmed), y=newConfirmed/Confirmed")
```
### log10(newConfirmed) ~ log10(Confirmed)

```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10') +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  ggtitle("x=Confirmed, y=newConfirmed")
```


## newConfirmed/yesterdayConfirmed
### newConfirmed/yesterdayConfirmed ~ Stamp
```{r}
df %>%
  mutate(yesterdayConfirmed=dplyr::lag(newConfirmed)) %>%
  ggplot(aes(x=as.POSIXct(Date), y=newConfirmed/yesterdayConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.75, se=FALSE, size=.33) +
  scale_x_datetime(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous(limits=c(0,3)) +
  ggtitle("x=Date, y=newConfirmed/yesterdayConfirmed")
```


### newConfirmed/yesterdayConfirmed ~ Confirmed
```{r}
df %>%
  mutate(yesterdayConfirmed=dplyr::lag(newConfirmed)) %>%
  ggplot(aes(x=Confirmed, y=newConfirmed/yesterdayConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.75, se=FALSE, size=.33) +
  scale_x_continuous(breaks=0:100*1000) +
  scale_y_continuous(limits=c(0,3)) +
  ggtitle("x=Date, y=newConfirmed/yesterdayConfirmed")
```


### newConfirmed/yesterdayConfirmed ~ log10(Confirmed)
```{r}
df %>%
  mutate(yesterdayConfirmed=dplyr::lag(newConfirmed)) %>%
  ggplot(aes(x=Confirmed, y=newConfirmed/yesterdayConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.75, se=FALSE, size=.33) +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,3)) +
  ggtitle("x=Date, y=newConfirmed/yesterdayConfirmed")
```



# Confirmed and Deaths growth

## Align growth curves from origin
Idea here is to compare the growth of Confirmed and Death counts from their respective beginning

- First 'Death' occurs 20 days after first 'Confirmed'
- Shift the 'Deaths' curve 20 days to left ('lagDeaths') and compare it with 'Confirmed'

```{r}
dd <- df %>%
  dplyr::select(Date, Confirmed, Deaths) %>%
  dplyr::mutate(lagDeaths=lead(Deaths,20))  %>%
  dplyr::filter(Date<as.POSIXct("2020-03-16"))
```
  
  
###  Confirmed+lagDeaths ~ Stamp (shifted)
```{r}
dd %>%
  tidyr::gather(key=Status, val=Count, Confirmed, lagDeaths) %>%
  ggplot(aes(x=Date, y=Count, colour=Status)) +
  scale_x_date(date_breaks = "2 day", date_labels="%d.%m") +
  scale_y_continuous(trans = 'log10', limits=c(1,1000)) +
  geom_point() + geom_line() +
  stat_smooth(span=0.85, se=FALSE, size=.33) +
  ggtitle("x=Date (shifted 20d), y=Confirmed,lagDeaths")
```

### lagDeaths/Confirmed ~ Stamp
```{r}
dd %>%
  ggplot(aes(x=Date, y=lagDeaths/Confirmed)) +
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.85, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "2 day", date_labels="%d.%m") +
  scale_y_continuous(breaks=seq(0,1.5,by=.25), limits=c(0,1.5)) +
   ggtitle("x=Date, y=lagDeaths/Confirmed")
```


## Align growth cureves from n=10
Idea here is again to compare the increase of Confirmed and Death counts

- Start from 10 cases
- Compare increase

```{r}
nRoot=10
rootDeaths <- df$Deaths[!is.na(df$Deaths) & df$Deaths>=nRoot]
rootConfirmed <- df$Confirmed[df$Confirmed>=nRoot][1:length(rootDeaths)]
```

### log10(Deaths) ~ log10(Confirmed) aligned on n=10
```{r}
dr <- data.frame(rConfirmed=rootConfirmed, rDeaths=rootDeaths)
rlm <- MASS::rlm(y=log10(dr$rDeaths),x=log10(dr$rConfirmed))
summary(rlm)
ggplot(data=dr, aes(x=rConfirmed,y=rDeaths)) + 
  geom_point() + geom_line(linetype=3) +
  scale_x_continuous(trans = 'log10', breaks=c(10,20,50,100,200,500,1000,2000,5000,10000,20000)) +
  scale_y_continuous(trans = 'log10', limits=c(10,NA), breaks=c(10,20,50,100,200,500,1000)) +
  stat_smooth(span=0.5, se=FALSE, size=.5) +
  ggtitle("")
```
  

### log10(Deaths) ~ log10(Confirmed )

```{r}
dal <- df %>% dplyr::filter(Deaths>1)
rlm <- MASS::rlm(x=log10(dal$Deaths),y=log10(dal$Confirmed))
ggplot(data=df%>% dplyr::filter(Deaths>1), aes(x=Confirmed, y=Deaths)) + 
  geom_point() + geom_line(linetype=3) + 
  stat_smooth(method=MASS::rlm, fullrange=TRUE, se=TRUE, size=.25) +
  scale_x_continuous(trans = 'log10', breaks=c(1000,2000,5000,10000),limits=c(1000,NA)) +
  scale_y_continuous(trans = 'log10') +
  ggtitle("x=Confirmed, y=Deaths")
```



# Hospitalisierung, Intensivstation
### Intensivstation/Hospitalisierung ~ Hospitalisierung

```{r}
ggplot(data=df%>% dplyr::filter(Hospitalisierung>0), aes(x=Hospitalisierung, y=Intensivstation/Hospitalisierung)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.8, se=FALSE, size=.25) +
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,NA)) +
  ggtitle("x=Hospitalisierung, y=Intensivstation/Hospitalisierung")
```

### newDeaths/Intensivstation ~ Stamp

```{r}
ggplot(data=df%>% dplyr::filter(Intensivstation>0), aes(x=Date, y=newDeaths/Intensivstation)) + 
  geom_point() + geom_line(linetype=3) + 
  stat_smooth(span=0.8, se=FALSE, size=.25) +
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,NA)) +
  ggtitle("x=Date, y=newDeaths/Intensivstation")
```


### newDeaths/Intensivstation ~ Intensivstation
```{r}
ggplot(data=df%>% dplyr::filter(Intensivstation>0), aes(x=Intensivstation, y=newDeaths/Intensivstation)) + 
  geom_point() + geom_line(linetype=3) + 
  stat_smooth(span=0.8, se=FALSE, size=.25) +
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,NA)) +
  ggtitle("x=Intensivstation, y=newDeaths/Intensivstation")
```

\newpage

# Derivatives
## LogScale
### logConfirmed~Date, span=.33
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.33
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

### logConfirmed~Date, span=.25
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.25
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

### logConfirmed~Date, span=.20
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.20
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

### logConfirmed~Date, span=.15
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.15
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

