---
title: "COVID-19-Austria: Evidence of spread slowing down"
author: "Thomas.Strehl@at.ibm.com"
date: "Mai 03, 2020"
geometry: "left=1.5cm,right=1cm,top=1.5cm,bottom=1.5cm"
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    fig_width: 7
    fig_height: 4
    fig_caption: false
    number_sections: true      
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, echo=FALSE}
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(zoo)
```

\newpage

# Corections to data
```{r}
# Several reported 'Tested' records seem to be wrong. 
# Here obvious errors are promoted to imputation
err <- data.frame(rbind(
c("B","2020-05-16","Tested",NA),
c("B","2020-05-31","Tested",NA),
c("B","2020-06-01","Tested",NA),
c("B","2020-06-21","Tested",NA),
c("B","2020-06-26","Tested",NA),
c("K","2020-06-21","Tested",NA),
c("Szb","2020-05-06","Tested",NA),
c("Szb","2020-06-09","Tested",NA),
c("Szb","2020-06-10","Tested",NA),
c("W","2020-04-15","Tested",32660),
c("W","2020-04-22","Tested",NA),
c("W","2020-04-23","Tested",NA),
c("W","2020-08-14","Tested",NA),
c("W","2020-08-15","Tested",NA),
c("W","2020-07-13","Tested",NA)
), stringsAsFactors=FALSE)
colnames(err)<-c("County","Date","Status","Value")
err$Value <- as.integer(err$Value)
```

# Status 
## Prepare data
```{r}
#csvFile <- "/home/at062084/DataEngineering/COVID-19/COVID-19-Austria/bmsgpk/data/COVID-19-austria.states.curated.status.csv"
csvFile <- "/home/at062084/DataEngineering/COVID-19/COVID-19-Austria/bmsgpk/data/COVID-19-austria.reconstructed.csv"
dv <- read.csv(csvFile, stringsAsFactors=FALSE) %>% 
  dplyr::mutate(Stamp=as.POSIXct(Stamp), Date=date(Stamp)) 

# grid of dates from first day
gridDate <- seq(ymd("2020-02-24"),ymd(format(now(),"%Y-%m-%d")), by='day')
gridStati=c("Tested", "Confirmed","Recovered","Deaths","Hospitalisierung","Intensivstation")
gridStatus <- rep(gridStati,length(gridDate))
gridDateStatus <- data.frame(Date=rep(gridDate,each=length(gridStati)), Status=rep(gridStati,times=length(gridDate)), stringsAsFactors=FALSE)

dv <- gridDateStatus %>% dplyr::left_join(dv, by=c("Date","Status"))

# Bug fix for wrong 'Tested' data from 14.8.2020 to 4.9.2020
#idxStamp <- dv$Stamp>=strptime("2020-08-14", format="%Y-%m-%d") & dv$Stamp<=strptime("2020-09-05", #format="%Y-%m-%d")
#idxStatus <- dv$Status=="Tested"
#idx <- idxStamp & idxStatus
#idx[is.na(idx)]<- FALSE
#dv[idx,"AT"] <- dv[idx,"AT"]*1000

# Correct 'Tested' data along err
#for(k in 1:dim(err)[1]) {
#  dv[dv$Date==strptime(err[k,"Date"],format="%Y-%m-%d") & dv$Status==err[k,"Status"] ,err[k,"County"]] <- #err[k,"Value"] 
#}

# All Counties
db <- dv %>% 
  dplyr::arrange(Stamp) %>%
  dplyr::group_by(Date,Status) %>%
  dplyr::filter(row_number()==n()) %>%
  dplyr::ungroup() %>% dplyr::ungroup() %>%
  dplyr::select(-Stamp) %>%
  tidyr::gather(key=County, value=Count, c(-Status,-Date)) %>%
  tidyr::spread(key=Status, val=Count) %>%
  dplyr::arrange(County,Date) %>%
  dplyr::group_by(County, add=FALSE) %>% 
  # impute missing values. Was difficult to get the syntax working !
  dplyr::mutate_at(vars(-group_cols()), imputeTS::na_interpolation, option="linear", maxgap=3) %>% 
  dplyr::ungroup() %>%
  dplyr::mutate(newConfirmed=Confirmed-lag(Confirmed)) %>%
  dplyr::mutate(newDeaths=Deaths-lag(Deaths)) %>%
  dplyr::mutate(newTested=Tested-lag(Tested)) %>%
  dplyr::mutate(newRecovered=Recovered-lag(Recovered)) %>%
  dplyr::select(1,2, 1+c(10,7, 8,2, 11,6, 9,3, 4,5))
tail(db,12)

# apply rolling mean to 'new*' cols
drm <- db %>%
  dplyr::arrange(Date, County) %>%
  dplyr::group_by(County) %>%
  dplyr::mutate_at(vars(starts_with("new")), rollmean, k=7, fill=NA, align="right") %>%
  dplyr::ungroup()

# KW's
dkw <- db %>% 
  dplyr::mutate(KW=week(Date)) %>% 
  dplyr::select(-Date) %>%
  dplyr::group_by(County, KW) %>%
  dplyr::summarise_all(sum) %>%
  dplyr::ungroup()

# AT
df <- db %>% 
  dplyr::filter(County=="AT") %>%
  dplyr::select(-County) 

dfrm <- drm %>% 
  dplyr::filter(County=="AT") %>%
  dplyr::select(-County) 

# Vienna:
dw <- db %>% 
  dplyr::filter(County=="W") %>%
  dplyr::select(-County) 

dwrm <- drm %>% 
  dplyr::filter(County=="W") %>%
  dplyr::select(-County) 

#dt <- ds %>% dplyr::filter(newConfirmed/newTested<.1, newTested>0)  

# newTested
ggplot(data=drm%>%dplyr::filter(County!="AT"), aes(x=Date, y=newTested, color=County)) + 
    geom_point(size=3, aes(shape=County)) +
    geom_line() +
    scale_shape_manual(values=c(1:10)) +
    #scale_x_continuous(limits=c(17,NA)) +
    scale_x_date(limits=c(as.Date(strptime("2020-05-01",format="%Y-%m-%d")),NA)) +
    scale_y_continuous(limits=c(0,10000)) +
    ggtitle("newTested ~ Date | County. Avg=7days")
  

#ggplot(data=drm%>%dplyr::filter(County!="AT"), aes(x=KW, y=newTested, color=County)) + 
#    geom_point(size=3, aes(shape=County)) +
#    geom_line() +
#    scale_shape_manual(values=c(1:10)) +
#    scale_x_continuous(limits=c(17,NA)) +
#    scale_y_continuous(limits=c(0,35000))

# newConfirmed
ggplot(data=drm%>%dplyr::filter(County!="AT"), aes(x=Date, y=newConfirmed, color=County)) + 
    geom_point(size=3, aes(shape=County)) +
    geom_line() +
    scale_shape_manual(values=c(1:10)) +
    #scale_x_continuous(limits=c(17,NA)) +
    scale_x_date(limits=c(as.Date(strptime("2020-05-01",format="%Y-%m-%d")),NA)) +
    scale_y_continuous(limits=c(.2,2000), trans="log10") +
    ggtitle("newConfirmed ~ Date | County. Avg=7days")

#ggplot(data=drm%>%dplyr::filter(County!="AT"), aes(x=KW, y=newConfirmed, color=County)) + 
#    geom_point(size=3, aes(shape=County)) +
#    geom_line() +
#    scale_shape_manual(values=c(1:10)) +
#    scale_x_continuous(limits=c(17,NA)) +
#    scale_y_continuous(limits=c(0,1000))

# newConfirmed/newTested
ggplot(data=drm, aes(x=Date, y=newConfirmed/newTested, color=County)) + 
  geom_point(size=3, aes(shape=County)) +
  geom_line() +
  scale_shape_manual(values=c(1:10)) +
  #scale_x_continuous(limits=c(17,NA)) +
  scale_x_date(limits=c(as.Date(strptime("2020-05-01",format="%Y-%m-%d")),NA)) +
  scale_y_continuous(limits=c(.001,.25), breaks=seq(0,.25,by=.01), trans="log10") +
  ggtitle("newConfirmed/newTested ~ Date | County. Avg=7days")

ggplot(data=drm %>% dplyr::filter(County %in% c("AT","B", "Ooe","W")), 
       aes(x=Date, y=newConfirmed/newTested, color=County)) + 
  geom_point(size=2, aes(shape=County)) +
  geom_line() +
  scale_shape_manual(values=c(1:10)) +
  #scale_x_continuous(limits=c(17,NA)) +
  scale_x_date(limits=c(as.Date(strptime("2020-05-01",format="%Y-%m-%d")),NA)) +
  scale_y_continuous(limits=c(0,.1), breaks=seq(0,.1,by=.01))+
  ggtitle("newConfirmed/newTested ~ Date | 3Counties. Avg=7days")

#ggplot(data=dt, aes(x=KW, y=newConfirmed/newTested, color=County)) + 
#  geom_point(size=3, aes(shape=County)) +
#  geom_line() +
#  scale_shape_manual(values=c(1:10)) +
#  scale_x_continuous(limits=c(17,NA)) +
#  scale_y_continuous(limits=c(0,.06), breaks=seq(0,.06,by=.01))

dc <- dw %>% 
  #dplyr::filter(!is.na(Tested)) %>% 
  dplyr::filter(month(Date)>=4) %>%
  dplyr::filter(newTested<20000) %>%
  dplyr::select(Date,newConfirmed,Confirmed,Tested,newTested)

ggplot(data=dc, aes(x=newTested, y=newConfirmed/newTested, shape=factor(month(Date)), color=month(Date))) +   geom_point(size=4) + 
  scale_y_continuous(limits=c(0,.075)) + 
  scale_color_gradient2(low = "red", mid="white", high="blue", midpoint=6.5) +
  scale_shape_manual(values=14+c(1:12)) + 
  scale_fill_brewer()

# Vienna tested and confirmed in one plot
scaled=1000
dwrmc <- dwrm %>% 
  dplyr::mutate(relConfirmedTested=newConfirmed/newTested*100, newTested=newTested/1000, newConfirmed_x10=newConfirmed/100) %>%
  dplyr::select(Date, relConfirmedTested, newTested, newConfirmed_x10) %>% 
  tidyr::gather(key=Status, val=Count, relConfirmedTested, newTested, newConfirmed_x10)

ggplot(data=dwrmc, aes(x=Date, y=Count, color=Status)) +
  scale_x_date(limits=c(as.Date(strptime("2020-07-01",format="%Y-%m-%d")),NA),
               date_breaks="1 weeks", date_labels="%a.%d.%m") +
  scale_y_continuous(limits=c(0,10), breaks=seq(0,10,by=1), 
                     sec.axis = sec_axis(~ . *scaled, breaks=seq(0,10000,by=1000))) +
  geom_point(size=1) + 
  geom_line() +
  ggtitle("newConfirmed,newTested ~ Date | Vienna. Avg=7days")


scaled=1000
dwrmr <- db %>% dplyr::filter(County=="W") %>% 
  dplyr::mutate(relConfirmedTested=newConfirmed/newTested*100, newTested=newTested/1000, newConfirmed_x10=newConfirmed/100) %>%
  dplyr::select(Date, relConfirmedTested, newTested, newConfirmed_x10) %>% 
  tidyr::gather(key=Status, val=Count, relConfirmedTested, newTested, newConfirmed_x10)

ggplot(data=dwrmr, aes(x=Date, y=Count, color=Status)) +
  scale_x_date(limits=c(as.Date(strptime("2020-07-01",format="%Y-%m-%d")),NA), 
               date_breaks="1 weeks", date_labels="%a.%d.%m") +
  scale_y_continuous(limits=c(0,10), breaks=seq(0,18,by=1), 
                     sec.axis = sec_axis(~ . *scaled, breaks=seq(0,18000,by=1000))) +
  geom_point(size=2) + 
  geom_line() 

```


## Status ~ Stamp

### Status ~ Stamp (all)

```{r}
ggplot(data=df %>% tidyr::gather(key=Status,val=Count, 2:8), aes(x=Date, y=Count, colour=Status, shape=Status)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous(trans = 'log10') +
  ggtitle("x=Date, y=log10(Count)")
```

### Status ~ Stamp (selected)

```{r}
ggplot(data=df %>% tidyr::gather(key=Status,val=Count, 2:8) %>% dplyr::filter(Status!="Tested", Status !="Confirmed", Status!="Recovered"), 
       aes(x=Date, y=Count, colour=Status, shape=Status)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous() +
  ggtitle("x=Date, y=Count")
```

### Confirmed ~ Stamp
```{r}
ggplot(data=df, aes(x=Date, y=Confirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.33, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous(breaks=0:13*1000) +
  ggtitle("x=Date, y=Confirmed")
```

## Tested
### Confirmed/Tested ~ Tested

```{r}
ggplot(data=df, aes(x=Tested, y=Confirmed/Tested)) + 
  geom_point() + geom_line(linetype=3) + 
  scale_x_continuous(breaks=seq(0,300000,by=20000)) +
#  scale_y_continuous(trans = 'log10') +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  ggtitle("x=Tested, y=Confirmed/Tested")
```

## newConfirmed
### newConfirmed ~ Stamp
```{r}
ggplot(data=df, aes(x=Date, y=newConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.25, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  ggtitle("x=Date, y=newConfirmed")
```

### newConfirmed ~ Confirmed
```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  scale_x_continuous(breaks=0:100*1000) +
  stat_smooth(span=0.33, se=FALSE, size=.33) +
  ggtitle("x=Confirmed, y=newConfirmed")
```

### newConfirmed ~ log10(Confirmed)
```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed)) + 
  scale_x_continuous(trans = 'log10', breaks=c(10,20,50,100,200,500,1000,2000,5000,10000)) +
#  scale_y_continuous(trans = 'log10') +
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.33, se=FALSE, size=.5) +
  ggtitle("x=log10(Confirmed), y=newConfirmed")
```



## newConfirmed/Confirmed
### newConfirmed/Confirmed ~ Date
```{r}
ggplot(data=df, aes(x=Date, y=newConfirmed/Confirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "1 week", date_labels="%d.%m") +
  ggtitle("x=Date, y=newConfirmed/Confirmed")
```

### newConfirmed/Confirmed ~ Confirmed
```{r}
ggplot(data=df, aes(x=(Confirmed), y=newConfirmed/Confirmed)) + 
  geom_point() + geom_line(linetype=3) +
   stat_smooth(span=0.5, se=FALSE, size=.33) +
  scale_x_continuous(breaks=0:100*1000) +
  ggtitle("x=Confirmed, y=newConfirmed/Confirmed")
```

### newConfirmed/Confirmed ~ log10(Confirmed)
```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed/Confirmed)) + 
  scale_x_continuous(trans = 'log10') +
  #scale_y_continuous(trans = 'log10') +
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  ggtitle("x=log10(Confirmed), y=newConfirmed/Confirmed")
```
### log10(newConfirmed) ~ log10(Confirmed)

```{r}
ggplot(data=df, aes(x=Confirmed, y=newConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10') +
  stat_smooth(span=0.5, se=FALSE, size=.33) +
  ggtitle("x=Confirmed, y=newConfirmed")
```


## newConfirmed/yesterdayConfirmed
### newConfirmed/yesterdayConfirmed ~ Stamp
```{r}
df %>%
  mutate(yesterdayConfirmed=dplyr::lag(newConfirmed)) %>%
  ggplot(aes(x=as.POSIXct(Date), y=newConfirmed/yesterdayConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.75, se=FALSE, size=.33) +
  scale_x_datetime(date_breaks = "1 week", date_labels="%d.%m") +
  scale_y_continuous(limits=c(0,3)) +
  ggtitle("x=Date, y=newConfirmed/yesterdayConfirmed")
```


### newConfirmed/yesterdayConfirmed ~ Confirmed
```{r}
df %>%
  mutate(yesterdayConfirmed=dplyr::lag(newConfirmed)) %>%
  ggplot(aes(x=Confirmed, y=newConfirmed/yesterdayConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.75, se=FALSE, size=.33) +
  scale_x_continuous(breaks=0:100*1000) +
  scale_y_continuous(limits=c(0,3)) +
  ggtitle("x=Date, y=newConfirmed/yesterdayConfirmed")
```


### newConfirmed/yesterdayConfirmed ~ log10(Confirmed)
```{r}
df %>%
  mutate(yesterdayConfirmed=dplyr::lag(newConfirmed)) %>%
  ggplot(aes(x=Confirmed, y=newConfirmed/yesterdayConfirmed)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.75, se=FALSE, size=.33) +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,3)) +
  ggtitle("x=Date, y=newConfirmed/yesterdayConfirmed")
```



# Confirmed and Deaths growth

## Align growth curves from origin
Idea here is to compare the growth of Confirmed and Death counts from their respective beginning

- First 'Death' occurs 20 days after first 'Confirmed'
- Shift the 'Deaths' curve 20 days to left ('lagDeaths') and compare it with 'Confirmed'

```{r}
dd <- df %>%
  dplyr::select(Date, Confirmed, Deaths) %>%
  dplyr::mutate(lagDeaths=lead(Deaths,20))  %>%
  dplyr::filter(Date<as.POSIXct("2020-03-16"))
```
  
  
###  Confirmed+lagDeaths ~ Stamp (shifted)
```{r}
dd %>%
  tidyr::gather(key=Status, val=Count, Confirmed, lagDeaths) %>%
  ggplot(aes(x=Date, y=Count, colour=Status)) +
  scale_x_date(date_breaks = "2 day", date_labels="%d.%m") +
  scale_y_continuous(trans = 'log10', limits=c(1,1000)) +
  geom_point() + geom_line() +
  stat_smooth(span=0.85, se=FALSE, size=.33) +
  ggtitle("x=Date (shifted 20d), y=Confirmed,lagDeaths")
```

### lagDeaths/Confirmed ~ Stamp
```{r}
dd %>%
  ggplot(aes(x=Date, y=lagDeaths/Confirmed)) +
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.85, se=FALSE, size=.33) +
  scale_x_date(date_breaks = "2 day", date_labels="%d.%m") +
  scale_y_continuous(breaks=seq(0,1.5,by=.25), limits=c(0,1.5)) +
   ggtitle("x=Date, y=lagDeaths/Confirmed")
```


## Align growth cureves from n=10
Idea here is again to compare the increase of Confirmed and Death counts

- Start from 10 cases
- Compare increase

```{r}
nRoot=10
rootDeaths <- df$Deaths[!is.na(df$Deaths) & df$Deaths>=nRoot]
rootConfirmed <- df$Confirmed[df$Confirmed>=nRoot][1:length(rootDeaths)]
```

### log10(Deaths) ~ log10(Confirmed) aligned on n=10
```{r}
dr <- data.frame(rConfirmed=rootConfirmed, rDeaths=rootDeaths)
rlm <- MASS::rlm(y=log10(dr$rDeaths),x=log10(dr$rConfirmed))
summary(rlm)
ggplot(data=dr, aes(x=rConfirmed,y=rDeaths)) + 
  geom_point() + geom_line(linetype=3) +
  scale_x_continuous(trans = 'log10', breaks=c(10,20,50,100,200,500,1000,2000,5000,10000,20000)) +
  scale_y_continuous(trans = 'log10', limits=c(10,NA), breaks=c(10,20,50,100,200,500,1000)) +
  stat_smooth(span=0.5, se=FALSE, size=.5) +
  ggtitle("")
```
  

### log10(Deaths) ~ log10(Confirmed )

```{r}
dal <- df %>% dplyr::filter(Deaths>1)
rlm <- MASS::rlm(x=log10(dal$Deaths),y=log10(dal$Confirmed))
ggplot(data=df%>% dplyr::filter(Deaths>1), aes(x=Confirmed, y=Deaths)) + 
  geom_point() + geom_line(linetype=3) + 
  stat_smooth(method=MASS::rlm, fullrange=TRUE, se=TRUE, size=.25) +
  scale_x_continuous(trans = 'log10', breaks=c(1000,2000,5000,10000),limits=c(1000,NA)) +
  scale_y_continuous(trans = 'log10') +
  ggtitle("x=Confirmed, y=Deaths")
```



# Hospitalisierung, Intensivstation
### Intensivstation/Hospitalisierung ~ Hospitalisierung

```{r}
ggplot(data=df%>% dplyr::filter(Hospitalisierung>0), aes(x=Hospitalisierung, y=Intensivstation/Hospitalisierung)) + 
  geom_point() + geom_line(linetype=3) +
  stat_smooth(span=0.8, se=FALSE, size=.25) +
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,NA)) +
  ggtitle("x=Hospitalisierung, y=Intensivstation/Hospitalisierung")
```

### newDeaths/Intensivstation ~ Stamp

```{r}
ggplot(data=df%>% dplyr::filter(Intensivstation>0), aes(x=Date, y=newDeaths/Intensivstation)) + 
  geom_point() + geom_line(linetype=3) + 
  stat_smooth(span=0.8, se=FALSE, size=.25) +
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,NA)) +
  ggtitle("x=Date, y=newDeaths/Intensivstation")
```


### newDeaths/Intensivstation ~ Intensivstation
```{r}
ggplot(data=df%>% dplyr::filter(Intensivstation>0), aes(x=Intensivstation, y=newDeaths/Intensivstation)) + 
  geom_point() + geom_line(linetype=3) + 
  stat_smooth(span=0.8, se=FALSE, size=.25) +
  #scale_x_continuous(trans = 'log10') +
  scale_y_continuous(limits=c(0,NA)) +
  ggtitle("x=Intensivstation, y=newDeaths/Intensivstation")
```

\newpage

# Derivatives
## LogScale
### logConfirmed~Date, span=.33
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.33
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

### logConfirmed~Date, span=.25
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.25
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

### logConfirmed~Date, span=.20
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.20
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

### logConfirmed~Date, span=.15
```{r}
dd <- dw %>% 
  dplyr::mutate(logConfirmed0=log(Confirmed),
                logConfirmed1=logConfirmed0-lag(logConfirmed0),
                logConfirmed2=logConfirmed1-lag(logConfirmed1)) %>%
  dplyr::select(Date,Confirmed, logConfirmed0,logConfirmed1, logConfirmed2) %>%
  dplyr::arrange(Date) %>%
  dplyr::mutate(nID=1:n())

span=.15
lss <- loess(logConfirmed1~nID, data=dd, span=span)

de <- dd %>%
  dplyr::mutate(logConfirmed1s = c(NA,lss$fitted)) %>%
  dplyr::mutate(logConfirmed2s=logConfirmed1s-lag(logConfirmed1s)) %>%
  dplyr::filter(!is.na(logConfirmed2))

ds <- de %>%
  dplyr::mutate(logConfirmed_20=logConfirmed0/20) %>%
  dplyr::select(Date, logConfirmed_20, logConfirmed1s,logConfirmed2s) %>%
  tidyr::gather(key=Derivative, value=Value,logConfirmed_20,logConfirmed1s,logConfirmed2s)

ggplot(data=ds, aes(x=Date, y=Value, colour=Derivative)) + geom_line() +
    scale_x_date(date_breaks = "1 week", date_labels="%d.%m") + 
    ggtitle(paste0("Derivatives of loess(logConfirmed, span=",span,"):  d/dt(logConfirmed~Date"))


```

